Functional assembly into and among community of a tropical evergreen forest

### 2016-08-05 inst/doc/draft.Rmd
v0.1.9028 Intro + Schema discussion

---
title: Functional assembly into and among community of a tropical evergreen forest from Western Ghats, India
author: "Sylvain Schmitt sylvain.schmitt@agroparistech.fr"
output: pdf_document
bibliography: refs.bib
---

@@

@@

Communities are composed from species spanning a range of functional traits responsible for their growth, survival and reproduction. Functional traits can thus be linked to trade-offs representative of species ecological strategies. For instance, specific leaf area varies along a soil fertility - shade index representative of a trade-off between ressource acquisition by photosynthesis and leaf investment in defense and leaf lifespan [@hodgson_is_2011]. Due to common ecological strategies, functional traits will covary in so called economic spectra. @wright_worldwide_2004 discovered the leaf economic spectrum followed by the wood economic spectrum of @chave_towards_2009. The leaf economic spectrum opposes productive ecological strategy to conservative persistent species. The wood economic spectrum is separating species between three pole close to @grime_evidence_1977 primary strategy axes: (i) competitive species, with high conductive efficiency, (ii) stress tolerant species, resistant to embolism, and (iii) disturbance tolerant species, with high mechanical strength. A debated question is links between species economic spectra and strategy axes. Whereas @baraloto_decoupled_2010 highlighted decoupled economic spectra allowing tree species to adjust differently their ecological strategies; @reich_world-wide_2014 advocate more for a whole plant economic spectrum identified by @freschet_evidence_2010 opposing fast growing species to more conservative ones.

Stochastic process and environmental filtering will be highly scale dependent. Effectively some functional traits will be more variable at the individual level with plasticity to micro-environmental conditions whereas others will be more affected on global environmental gradient with phylogeny. For instance, @messier_trait_2016 argued the scale to which traits are covarying in ecological strategy, more especially they question the existence of the leaf economic spectrum at community scale.

+ *Move to India Western Ghats context*
    - *No previous study of community ecology with functional traits*
    - *Peculiar case of highly seasonal moist evergreen forest*
+ *Move to previous questionning*
+ *Quid from ecological strategies at small scale in Uppangala forest*
+ *Quid from partitionning between stochastic event and environmental filtering on community structure*

@@

```{r 1 Traits}
rm(list = ls()); invisible(gc())
PFT <- genPFT()
table <- data.frame(
  traits = c('Thickness', 'Leaf area', 'Leaf dry matter content', 'Specific leaf area', 'Wood density'),
  abbreviation = c('Thick', 'LA', 'LDMC', 'SLA', 'WD'),
  range = c(paste(range(PFT$Thick, na.rm = T), collapse = '-'),
            paste(round(range(PFT$LA, na.rm = T)), collapse = '-'),
            paste(round(range(PFT$LDMC, na.rm = T)), collapse = '-'),
            paste(round(range(PFT$SLA, na.rm = T), 2), collapse = '-'),
            paste(round(range(PFT$WD, na.rm = T), 2), collapse = '-')),
  `mean (sd)` = c(paste0(round(mean(PFT$Thick, na.rm = T)), ' (',
                         round(sd(PFT$Thick, na.rm = T)), ')'),
                  paste0(round(mean(PFT$LA, na.rm = T)), ' (',
                         round(sd(PFT$LA, na.rm = T)), ')'),
                  paste0(round(mean(PFT$LDMC, na.rm = T)), ' (',
                         round(sd(PFT$LDMC, na.rm = T)), ')'),
                  paste0(round(mean(PFT$SLA, na.rm = T),2), ' (',
                         round(sd(PFT$SLA, na.rm = T),2), ')'),
                  paste0(round(mean(PFT$WD, na.rm = T),2), ' (',
                         round(sd(PFT$WD, na.rm = T),2), ')')),
  unit = c('$\\mu m$', '$mm^2$', '$mg.g^-1$', '$m^2.kg^-1$', '$g.cm^-3$'),
  strategy = c('Leaf defense vs. investment',
               'Leaf ressource capture vs.investment',
               'Leaf defense vs. investment',
               'Leaf ressource acquisiton vs. defense',
               'Stem transport, structure and defense')
  )
kable(table,  label="traits", row.names = F)
rm(table)
```

@@

To investigate local functional composition of communities we used community weighted means (CWM), community weighted variances (CWV), and coefficients of variation from the nearest neighbour distance (CVNND) [CWM and CWV, sensu @violle_let_2007, CVNND, @jung_intraspecific_2010]. CWM for each trait $t$ was calculated as the mean of species trait values, $t_{i}$, of the $S$ species in each community, with each species, $i$, weighted by its relative abundance, $p_{i}$:
$$CWM = \sum_{i=1}^{S} p_{i} * t_{i} ~ (1)$$
CWM was also computed in presence-absence, to further emphasize role of rare species in community trait distribution. Linear regressions between CWMs and biotic and abiotic factors ($CWM_{trait} = Slope + Curvature + Wetness + SW + Canopy + BA$), after stepwise selection, were compared to corresponding 999 null linear regressions from NM1 with F-statistic.

@@

All analysis were conducted using the R 3.3.0  software environment [@r_core_team_r:_2016] with packages 
`ade4` [@dray_ade4_2007], 
`ape` [@paradis_ape:_2004], 
`phylosignal` [@keck_phylosignal:_2015], 
`spacodiR` [@eastman_spacodir:_2013],
`vegan` [@oksanen_vegan:_2016].

@@

Community weighted means (CWM) only show few significant trends along environmental gradients (see table 3). CWMs reveal that local mean WD will significantly increase under South-West; whereas SLA and LA, in presence-absence only, tend to decrease with curvature. In addition, LDMC in presence-absence is decreasing with canopy height.

Table 3: Summary of linear regressions and null models for CWMs of traits. _Linear regression results for each trait CWMs by species relative abundance and presence after a variable stepwise selection with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA). Linear regressions global and partial F-statistics were compared to 999 null linear regressions built with null model NM1 which randomizes traits among species. See table 1 for abbreviations. The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._

@@

# Result tables
kable(bold(regression_table(CWM.lm[[1]], reg, pval[,1], ppval[[1]])))
kable(bold(regression_table(CWM.lm[[2]], reg, pval[,2], ppval[[2]])))
rm(CWM.lm, nCWM.lm, pval, ppval, reg) ; invisible(gc())
```

Traits divergence is measured with community weighted variances (CWV). Overall, observed CWVs doesn't differ from null expectations for all traits at the exception of one community for WD (over 240 communities). Still we can observe some trends between standardized effect size (SES) of CWVs and environment (see table 4). Effectively South-West orientation will make local functional composition of communities converge towards a mean LES and LDMC whereas LA will diverge inside community under South-West orientation. LA and LDMC values will also diverge with increasing curvature. Observed CVNNDs doesn't reveal niche differenciation among communities for all traits. Still we can observe some trends between standardized effect size (SES) of CVNNDs and environment (see table 4). Species tend to differentiate their LA with lower curvature and wetness and their WD opppositely to SouthWest orientation.

Table 4: Summary of linear regressions for traits standardized effect size (SES) of CWVs and CVNNDs. _Linear regression results for each trait SES of CWVs and CVNNDs after a variablestepwise selection with slope, curvature, wetness, and southwesterness (SW), canopy height (Canopy) and basal area (BA). See table 1 for abbreviations. The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._

@@

# Results
# barplot(t(CWV.pval.res), las = 2, col = c('darkblue', 'firebrick'), ylim = c(0,1), ylab = 'Proportion of non null communities') ; legend('topright', c('Inferior','Superior'), fill = c('darkblue', 'firebrick'))
kable(bold(CWV.ses.res))
rm(CWV.pval, CWV.pval.res, CWV.ses.res, CWV.ses.lm, nCWV)
```

@@

# Results
# barplot(t(CVNND.pval.res), las = 2, col = c('darkblue'), ylim = c(0,1), ylab = 'Proportion of non null communities') ; legend('topright', c('Inferior'), fill = c('darkblue'))
kable(bold(CVNND.ses.res))
rm(CVNND.pval, CVNND.pval.res, CVNND.ses.res, CVNND.ses.lm, nCVNND) ; invisible(gc())
```

@@

Whereas LDMC is not phylogenetically convergent nor divergent (Moran's I of 0 with a $p-value > 0.1$), LA and Thick are significantly convergent among the phylogenetic tree (Moran's I of 0.021 and 0.018 with a $p-value < 0.001$), and WD and SLA are weakly convergent (Moran's I of 0.005 and 0.003 with $p-value < 0.1$). Phylogenetic composition between communities is not correlated to environement spatial organization with Mantel test ($r_{\beta_{st},E} = 0.02$ and $r_{\pi_{st},E} = 0.05$ with $p-value > 0.1$) nor with spatial residuals represented with partial Mantel test knowing environment ($r_{\beta_{st},S|E} = 0.01$ and $r_{\pi_{st},S|E} = -0.01$ with $p-value > 0.1$ respectively).

@@

Plants distinguish theimselves in functional trait composition betwen individuals and even more between species. Variation of species functional composition results from functional trade-offs that compose major variation axes of functional traits called economic spectra. We found back the leaf economic spectrum of @wright_worldwide_2004 opposing in our case SLA to LDMC in a trade-off between productivity, with less carbon investment for higher photosynthesis rates, opposed to persistance, with more leaf defense and investment. Independantly to that spectrum, we detected a second major axis of trait variation opposing WD to LA and Thick. This axis opposes more resistant to stress and disturbance species, thanks to denser wood, to competitive species, with thicker and larger leaves for a more efficient water conductivity. This spectrum correspond to the wood economic spectrum described by @chave_towards_2009, and more precisely this opposition between small thin leave with dense wood to large thick leaves with light wood correspond to unexplained results of @baraloto_decoupled_2010.

@@

The leaf economic spectrum is majorly affected by aspect with a convergence of value toward optimal functional traits under South-West orientation, especially for rare species with LDMC. Aspect, and more specifically south-west, effect in tropical forest is scarcely documented but highlighted in discussion by @gunatilleke_specieshabitat_2006. @bunyan_effect_2015 observed a similar effect on shola distribution of Western Ghats and suggested it could result from monsoon winds. Effectively monsoon winds arise fom south-west direction in this region, and bring more stressfull and disturbed conditions to species [@soman_aspects_1990]. Curvature is also affecting the leaf economy. Convex area will accumulate soil and attract water flux increasing soil fertiliy conditions for species that will invest more in leaf ressource acquisition toward a productive strategy with higher SLA and a convergent optimal LDMC. SLA and LDMC with curvature effect seems to reflect well soil fertility [@hodgson_is_2011].

Biotic factors also affect the LES through the negative effect of canopy height on LDMC. The canopy height effect in fact is null on understorey species and only apply on species that take part in the canopy composition or emerge (results unshown). We could have assumed that understorey species would be investing in higher LDMC for an higher leaf lifespan under shade stress [@kitajima_tissue-level_2010]. But both the high variation of canopy height resulting in a relatively light open canopy and the dry season of 3.8 months suggest that leaf lifespan investment is not so obvious. Instead of a growth-survival trade-off shade tolerant species may adopt a more low-high light growth trade-off strategy [@poorter_leaf_2009] strongly affecting SLA but not driving LDMC. Emergent and high canopy species will have a lower LDMC, resulting from a more productive strategy toward leaf resource acquisition at the expense of leaf carbon investment in defense and resistance. Thus canopy species with the lowest LDMC will grow the fastest and then reach the highest canopy. Measurement of LDMC on species classified as canopy and emergent are realized most of the time on shadow leaves from understorey individuals. Still we observed previous relation indicating that species phylogenetic determinism is stronger than micro-environmental variation plasticity for LDMC, as suggested by the intra-specific variation of LDMC 1.5 times bigger than intra-individuals variation.

The wood economic spectrum defined by WD, LA, and Thick is only partially affected by environment at our 10 ha plot scale. Neutral hypothesis of funcional equivalence [@hubbell_neutral_2005] are most of the times sufficient to build back similar links between topography and functional compoisiton of communities. Still aspect wiht south-west orientation and curvature are the two main ecological factors filtering functional traits from the wood economic spectrum. South-west orientation, with probably more disturbance and stress, is globaly increasing community WD mean values while reducing eveness by convergence toward different optimal WD. In parrallel, LA is broaden with south-west direction, pehraps to answer to hydraulic diversity induced by WD multimodality, as WD and LA are mechanically linked [@coomes_scaling_2008]. Fertility decrease, with curvature increase, results in smaller leaves inside communities with a distribution broaden by multimodality reducing eveness. Wetness is increasing with higher curvature and lower LA eveness, indicating that multimodality is happening more in wettest low elevation area. Wetness could compensate water departure trigerred by positive curvature in lower elevation areas.

@@

Limited dispersal, assumed by neutral theory [@condit_beta-diversity_2002], does not affect florisitic composition of communities at the 10 ha plot scale due to the absence of links  with spatial residuals. Florisitic composition of communities is only shaped by environment. On opposite, functional composition is overall not structured with environment nor with spatial residuals.At the 10 ha level functional variation between communities is stochastic; whereas at the community level environment is partially shapping florisitic composition. Measured functional traits should vary to a larger scale [@messier_trait_2016]. Still WD is positively correlated to $\beta$ structure. **To developp**. WD of rare species is even diverging with spatial residuals supposing an unaccounted spatially organized variable opposing close community in WD composition. Topography removed from spatial residuals, and climate and pedology being constant in the 10 ha plot, this unaccouted variation is assumed to be due to stochastic event in plot history as death. Unshowed analysis reveals that this spatial divergence of WD is not explained by tree death events from the two last decades but certainly to older events. Finally, variation in distance between taxon wasn't translated in phylogenetic variation whereas most of measured functional traits appeared phylogenetically conserved and clustered.

@@

Ecological strategies underline two econmic spectra as shown by @baraloto_decoupled_2010. Whereas the multivariate analysis of our functional traits lack trait diversity for correct interpretations of functional trade-offs. Measured traits were all morphological and from the leaf except for one. Anatomical and chemical traits as leaf vein patterns and leaf chemical content (N, P, K) could brought broader perspective on ecological strategies of Uppangala species. Still the study reveal that few commonly used traits can highlight well known strategy axes.

Besides patterns have been detected between functional composition and environment at both local ($\alpha$) and between-community ($\beta$) scales. Lot of relation observed didn't differ from those obtain with null hypothesis reflecting neutral theory [as functional equivalence, @hubbell_neutral_2005]. Ecological gradient and spatial scale sizes can be at the origin of this lack of environmental filtering. Effectively the abiotic variation is mainly topographic with a constant climate and pedology. Study highlighting environmental filtering often present a higher environmental variation [@bernard-verdier_community_2012]. Similar to this, the 10 ha plot scale is really fine and partially responsible for the small environmental variation. @kraft_functional_2008 detected topographical filtering on functional trait and @gunatilleke_specieshabitat_2006 observed niche differentiation on 25 ha tropical forest plots including large flat valleys. This mid-low elevation study of functional traits from western ghats forest could be reproduced at higher elevation or in dryer areas as eastern Ghats for better insight of environmental filtering at broad scale. On the other hand WD as shown to be plastic to environment at fine scale both in $\alpha$ and $\beta$ levels. Species WD seems highly adptated to environment in western Gahts forests in link to monsoon events.

>>>>

---
title: Functional assembly into and among community of a tropical evergreen forest
  from Western Ghats, India
author: "Sylvain Schmitt sylvain.schmitt@agroparistech.fr"
date: '`r Sys.Date()`'
output: word_document
bibliography: refs.bib
---

@@

Built with `Uppangala` package version `r packageVersion('Uppangala')`.

@@

Communities are composed from species spanning a range of functional traits responsible for their growth, survival and reproduction. Functional traits can thus be linked to trade-offs representative of species ecological strategies. For instance, specific leaf area varies along a soil fertility - shade index representative of a trade-off between ressource acquisition by photosynthesis and leaf investment in defense and leaf lifespan [@hodgson_is_2011]. Due to common ecological strategies, functional traits will covary in so called economic spectra. @wright_worldwide_2004 discovered the leaf economic spectrum followed by the wood economic spectrum of @chave_towards_2009. The leaf economic spectrum opposes productive ecological strategy to conservative persistent species. The wood economic spectrum is separating species between three pole close to @grime_evidence_1977 primary strategy axes: (i) competitive species, with high conductive efficiency, (ii) stress tolerant species, resistant to embolism, and (iii) disturbance tolerant species, with high mechanical strength. A debated question is links between species economic spectra and strategy axes. Whereas @baraloto_decoupled_2010 highlighted decoupled economic spectra allowing tree species to adjust differently their ecological strategies; @reich_world-wide_2014 advocate more for a whole plant economic spectrum identified by @freschet_evidence_2010 opposing fast growing species to more conservative ones. Stochastic process and environmental filtering will be highly scale dependent. Effectively some functional traits will be more variable at the individual level with plasticity to micro-environmental conditions whereas others will be more affected on global environmental gradient with phylogeny. For instance, @messier_trait_2016 argued the scale to which traits are covarying in ecological strategy, more especially they question the existence of the leaf economic spectrum at community scale.

Functional traits in community ecology are linking ecological strategies, community assembly theory and functional diversity [@westoby_land-plant_2006]. In this paper we present the use of traits-based approach combined to community ecology to encompass functional assembly into and among community of a tropical evergreen forest from Western Ghats, India. Few studies investigated functional ecology whithin the Western Ghats indian biodiversity hotspot [@myers_biodiversity_2000]. Western Ghats present atypic moist tropical evergreen forest though seasonnaly dry and afffected by strong monsoon events. Such environmental conditions can be highly interesting to discuss and evaluate community assembly theory resulting in functional assembly of those forests. To do so we organised our study framework in two main questions :

@@

```{r 1 Traits}
rm(list = ls()[-which(ls() == "cores")]); invisible(gc())
PFT <- genPFT()
table <- data.frame(
  Trait = c('Thickness', 'Leaf area', 'Leaf dry matter content', 'Specific leaf area', 'Wood density'),
  Abbreviation = c('Thick', 'LA', 'LDMC', 'SLA', 'WD'),
  Range = c(paste(range(PFT$Thick, na.rm = T), collapse = '-'),
            paste(round(range(PFT$LA, na.rm = T)), collapse = '-'),
            paste(round(range(PFT$LDMC, na.rm = T)), collapse = '-'),
            paste(round(range(PFT$SLA, na.rm = T), 2), collapse = '-'),
            paste(round(range(PFT$WD, na.rm = T), 2), collapse = '-')),
  m = c(paste0(round(mean(PFT$Thick, na.rm = T)), ' (',
                         round(sd(PFT$Thick, na.rm = T)), ')'),
                  paste0(round(mean(PFT$LA, na.rm = T)), ' (',
                         round(sd(PFT$LA, na.rm = T)), ')'),
                  paste0(round(mean(PFT$LDMC, na.rm = T)), ' (',
                         round(sd(PFT$LDMC, na.rm = T)), ')'),
                  paste0(round(mean(PFT$SLA, na.rm = T),2), ' (',
                         round(sd(PFT$SLA, na.rm = T),2), ')'),
                  paste0(round(mean(PFT$WD, na.rm = T),2), ' (',
                         round(sd(PFT$WD, na.rm = T),2), ')')),
  Unit = c('$\\mu m$', '$mm^2$', '$mg.g^-1$', '$m^2.kg^-1$', '$g.cm^-3$'),
  s = c('Leaf defense vs. investment',
               'Leaf ressource capture vs.investment',
               'Leaf defense vs. investment',
               'Leaf ressource acquisiton vs. defense',
               'Stem transport, structure and defense')
  )
names(table)[c(4,6)] <- c('Mean (sd)', 'Trade-off')
kable(table, row.names = F)
rm(table)
```

@@

To investigate local functional composition of communities we used community weighted means (CWM), community weighted variances (CWV), and coefficients of variation from the nearest neighbour distance (CVNND) [CWM and CWV, sensu @violle_let_2007; CVNND, @jung_intraspecific_2010]. CWM for each trait $t$ was calculated as the mean of species trait values, $t_{i}$, of the $S$ species in each community, with each species, $i$, weighted by its relative abundance, $p_{i}$:
$$CWM = \sum_{i=1}^{S} p_{i} * t_{i} ~ (1)$$
CWM was also computed in presence-absence, to further emphasize role of rare species in community trait distribution. Linear regressions between CWMs and biotic and abiotic factors ($CWM_{trait} = Slope + Curvature + Wetness + SW + Canopy + BA$), after stepwise selection, were compared to corresponding 999 null linear regressions from NM1 with F-statistic.

@@

All analysis were conducted using the R 3.3.0  software environment [@r_core_team_r:_2016] with packages 
`ade4` [@dray_ade4_2007], 
`ape` [@paradis_ape:_2004], 
`phylosignal` [@keck_phylosignal:_2015], 
`spacodiR` [@eastman_spacodir:_2013],
and `vegan` [@oksanen_vegan:_2016].

@@

Community weighted means (CWM) only show few significant trends along environmental gradients (see table 3). CWMs reveal that local mean WD will significantly increase under south-west; whereas SLA and LA, in presence-absence only, tend to decrease with curvature. In addition, LDMC in presence-absence is decreasing with canopy height.

Table 3: Summary of linear regressions and null models for CWMs of traits. _Linear regression results for each trait CWMs weighted by species relative abundance (AB) and presence-absence (PA) after a variable stepwise selection with slope, curvature, wetness, southwesterness (SW), canopy height (Canopy) and basal area (BA). Linear regressions global and partial F-statistics were compared to 999 null linear regressions built with null model NM1 which randomizes traits among species. See table 1 for abbreviations. The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._

@@

# Result tables
t1 <- regression_table(CWM.lm[[1]], reg, pval[,1], ppval[[1]])
t1 <- cbind(c('AB', rep(' ', length(t1[,1])-1)), t1)
names(t1)[1] <- ' '
t2 <- regression_table(CWM.lm[[2]], reg, pval[,2], ppval[[2]])
t2 <- cbind(c('PA', rep(' ', length(t2[,2])-1)), t2)
names(t2)[1] <- ' '
kable(bold(rbind(t1,t2)), row.names = F)
rm(CWM.lm, nCWM.lm, pval, ppval, reg, t1, t2) ; invisible(gc())
```

Traits divergence is measured with community weighted variances (CWV). Overall, observed CWVs doesn't differ from null expectations for all traits at the exception of one community for WD (over 240 communities). Still we can observe some trends between standardized effect size (SES) of CWVs and environment (see table 4). Effectively south-west orientation will make local functional composition of communities converge towards a mean LES and LDMC whereas LA will diverge inside community under south-west orientation. LA and LDMC values will also diverge with increasing curvature. Observed CVNNDs doesn't reveal niche differenciation among communities for all traits. Still we can observe some trends between standardized effect size (SES) of CVNNDs and environment (see table 4). Species tend to differentiate their LA with lower curvature and wetness and their WD opppositely to SouthWest orientation.

Table 4: Summary of linear regressions for traits standardized effect size (SES) of CWVs and CVNNDs. _Linear regression results for each trait SES of CWVs and CVNNDs after a variable stepwise selection with slope, curvature, wetness southwesterness (SW), canopy height (Canopy) and basal area (BA). See table 1 for abbreviations. The number of asterisk indicate the p-value threshold ( * p-value<0.05, ** p-value <0.01, *** p-value<0.001)._

@@

# Results
# barplot(t(CWV.pval.res), las = 2, col = c('darkblue', 'firebrick'), ylim = c(0,1), ylab = 'Proportion of non null communities') ; legend('topright', c('Inferior','Superior'), fill = c('darkblue', 'firebrick'))
CWV.ses.res <- cbind(c('CWV', rep(' ', length(CWV.ses.res[,1])-1)), row.names(CWV.ses.res), CWV.ses.res)
names(CWV.ses.res)[1:2] <- c(' ', 'Trait')
rm(CWV.pval, CWV.pval.res, CWV.ses.lm, nCWV)
```

@@

# Results
# barplot(t(CVNND.pval.res), las = 2, col = c('darkblue'), ylim = c(0,1), ylab = 'Proportion of non null communities') ; legend('topright', c('Inferior'), fill = c('darkblue'))
CVNND.ses.res <- cbind(c('CVNND', rep(' ', length(CVNND.ses.res[,1])-1)), row.names(CVNND.ses.res), CVNND.ses.res)
names(CVNND.ses.res)[1:2] <- c(' ', 'Trait')
kable(bold(rbind(CWV.ses.res,CVNND.ses.res)), row.names = F)
rm(CVNND.pval, CVNND.pval.res, CVNND.ses.res, CVNND.ses.lm, nCVNND) ; invisible(gc())
```

@@

Whereas LDMC is not phylogenetically convergent nor divergent (Moran's I of 0 with a $p-value > 0.1$), LA and Thick are significantly convergent among the phylogenetic tree (Moran's I of 0.021 and 0.018 with a $p-value < 0.001$), and WD and SLA are weakly convergent (Moran's I of 0.005 and 0.003 with $p-value < 0.1$). Phylogenetic composition between communities is not correlated to environement spatial organization with Mantel test ($r_{\beta_{st},E} = 0.02$ and $r_{\pi_{st},E} = 0.05$ with $p-value > 0.1$) nor with spatial residuals represented with partial Mantel test knowing environment ($r_{\beta_{st},S|E} = 0.01$ and $r_{\pi_{st},S|E} = -0.01$ with $p-value > 0.1$).

@@

Plants distinguish theimselves in functional trait composition betwen individuals and even more between species. Variation of species functional composition results from functional trade-offs that compose major variation axes of functional traits called economic spectra. We found back the leaf economic spectrum of @wright_worldwide_2004 opposing in our case SLA to LDMC in a trade-off between productivity, with less carbon investment for higher photosynthesis rates, opposed to persistance, with more leaf defense and investment. Independantly to that spectrum, we detected a second major axis of trait variation opposing WD to LA and Thick. This axis opposes more resistant to stress and disturbance species, thanks to denser wood, to competitive species, with thicker and larger leaves for a more efficient water conductivity. This spectrum correspond to the wood economic spectrum described by @chave_towards_2009, and more precisely this opposition between small thin leave with dense wood to large thick leaves with light wood correspond to the additional gradient uncovered by  @baraloto_decoupled_2010.

@@

The leaf economic spectrum is majorly affected by aspect with a convergence of value toward optimal functional traits under south-west orientation, especially for LDMC of rare species. Aspect, and more specifically south-west, effect in tropical forest is scarcely documented but highlighted in discussion by @gunatilleke_specieshabitat_2006. @bunyan_effect_2015 observed a similar effect on shola distribution of Western Ghats and suggested it could result from monsoon winds. Effectively monsoon winds arise fom south-west direction in this region, and bring more stressfull and disturbed conditions to species [@soman_aspects_1990]. Curvature is also affecting the leaf economy. Convex area will accumulate soil and attract water flux increasing soil fertiliy conditions for species that will invest more in leaf ressource acquisition toward a productive strategy with higher SLA and a convergent optimal LDMC. SLA and LDMC with curvature effect seems to reflect well soil fertility [@hodgson_is_2011].

Biotic factors also affect the LES through the negative effect of canopy height on LDMC. The canopy height effect is null on understorey species and only apply on species that take part in the canopy composition or emerge (results unshown). We would assumed that understorey species would be investing in higher LDMC for an higher leaf lifespan under shade stress [@kitajima_tissue-level_2010]. But both the high variation of canopy height resulting in a relatively light open canopy and the dry season of 3.8 months suggest that leaf lifespan investment is not so obvious. Instead of a growth-survival trade-off shade tolerant species may adopt a more low-high light growth trade-off strategy [@poorter_leaf_2009] strongly affecting SLA but not driving LDMC. Emergent and high canopy species will have a lower LDMC, resulting from a more productive strategy toward leaf resource acquisition at the expense of leaf carbon investment in defense and resistance. Thus canopy species with the lowest LDMC will grow the fastest and then reach the highest canopy. Measurement of LDMC on species classified as canopy and emergent are realized most of the time on shadow leaves from understorey individuals. Still we observed previous relation indicating that species phylogenetic determinism is stronger than micro-environmental variation plasticity for LDMC, as suggested by the intra-specific variation of LDMC 1.5 times bigger than intra-individuals variation.

The wood economic spectrum defined by WD, LA, and Thick is only partially affected by environment at our 10 ha plot scale. Neutral hypothesis of funcional equivalence [@hubbell_neutral_2005] are most of the times sufficient to build back similar links between topography and functional compoisiton of communities. Still aspect wiht south-west orientation and curvature are the two main ecological factors filtering functional traits from the wood economic spectrum. South-west orientation, with probably more disturbance and stress from monsoon, result in denser wood with reduced eveness by convergence toward different optimal density. In parrallel, south-west direction broaden leaf size strategies, pehraps to answer to hydraulic diversity induced by WD multimodality, as WD and LA are mechanically linked [@coomes_scaling_2008]. Fertility decrease, with concavity, results in smaller leaves inside communities with a distribution broaden by multimodality reducing eveness. Wetness is increasing with higher curvature and lower LA eveness, indicating that multimodality is happening more in wettest low elevation area. Wetness could compensate water departure trigerred by positive curvature in lower elevation areas.

@@

Limited dispersal, assumed by neutral theory [@condit_beta-diversity_2002], does not affect florisitic composition of communities at the 10 ha plot scale revealed by the absence of links  with spatial residuals. Florisitic composition of communities is only shaped by environment. On opposite, functional composition is overall not structured with environment nor with spatial residuals. At the 10 ha level functional variation between communities is stochastic; whereas at the community level environment is partially shapping florisitic composition. Measured functional traits should vary to a larger scale [@messier_trait_2016]. Still environment is shaping community resistance to disturbance and stress by affecting wood density $\beta$ structure at fine scale. WD of rare species is even diverging with spatial residuals supposing an unaccounted spatially organized variable opposing close community in WD composition. Topography removed from spatial residuals, and climate and pedology being constant in the 10 ha plot, this unaccouted variation is assumed to be due to stochastic event in plot history as death. Unshowed analysis reveals that this spatial divergence of WD is not explained by tree death events from the two last decades but certainly to older events. Finally, variation in distance between taxon wasn't translated in phylogenetic variation whereas most of measured functional traits appeared phylogenetically conserved and clustered.

@@

Ecological strategies underline two econmic spectra as shown by @baraloto_decoupled_2010. Whereas the multivariate analysis of our functional traits lack trait diversity for correct interpretations of functional trade-offs. Measured traits were all morphological and from the leaf except for one. Anatomical and chemical traits as leaf vein patterns and leaf chemical content (N, P, K) could brought broader perspective on ecological strategies of Western Ghats species. Still the study reveal that few commonly used traits can highlight well known strategy axes.

Besides patterns have been detected between functional composition and environment at both local ($\alpha$) and between-community ($\beta$) scales. Lot of relation observed didn't differ from those obtain with null hypothesis reflecting neutral theory [as functional equivalence, @hubbell_neutral_2005]. Ecological gradient and spatial scale sizes can be at the origin of this lack of environmental filtering. Effectively the abiotic variation is mainly topographic with a constant climate and pedology. Study highlighting environmental filtering often present a higher environmental variation [@bernard-verdier_community_2012]. In parallel, the 10 ha plot scale is really low and partially responsible for the small environmental variation. @kraft_functional_2008 detected topographical filtering on functional trait and @gunatilleke_specieshabitat_2006 observed niche differentiation on 25 ha tropical forest plots including large flat valleys. This mid-low elevation study of functional traits from Western Ghats forest could be reproduced at higher elevation or in dryer areas as Eastern Ghats for better insight of environmental filtering at broad scale. On the other hand, WD as shown to be filtered by environment at fine scale both in $\alpha$ and $\beta$ levels. Species wood seems highly adpted to environment in Western Gahts forests in link to monsoon events.
